---
title: "Data Consistency Analysis"
subtitle: "Household Survey — Template"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    latex_engine: pdflatex
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
  - \floatplacement{table}{H}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhead[L]{Data Consistency Analysis}
  - \fancyhead[R]{Household Survey}
  - \fancyfoot[C]{\thepage}
---

```{r setup, include=FALSE}
packages <- c("readxl", "tidyverse", "knitr", "kableExtra")

for(p in packages) {
  if(!require(p, character.only = TRUE, quietly = TRUE)) {
    install.packages(p, quiet = TRUE)
    library(p, character.only = TRUE)
  }
}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.kable.NA = '-')
```

<!-- ================================================================== -->
<!-- LAYER 1: INCIDENCE ACCUMULATOR                                      -->
<!-- This list collects every finding as the document generates.         -->
<!-- At the end, the accumulated incidences determine the ruling.        -->
<!-- ================================================================== -->

```{r incidence-system}
incidences <- list(
  critical = data.frame(Section = character(), Variable = character(), 
                        Description = character(), N_Affected = integer(),
                        stringsAsFactors = FALSE),
  medium   = data.frame(Section = character(), Variable = character(), 
                        Description = character(), N_Affected = integer(),
                        stringsAsFactors = FALSE),
  minor    = data.frame(Section = character(), Variable = character(), 
                        Description = character(), N_Affected = integer(),
                        stringsAsFactors = FALSE)
)

add_incidence <- function(severity, section, variable, description, n = NA) {
  new_row <- data.frame(Section = section, Variable = variable, 
                        Description = description, N_Affected = n,
                        stringsAsFactors = FALSE)
  if(severity == "critical")    incidences$critical <<- rbind(incidences$critical, new_row)
  else if(severity == "medium") incidences$medium   <<- rbind(incidences$medium, new_row)
  else                          incidences$minor    <<- rbind(incidences$minor, new_row)
}
```

<!-- ================================================================== -->
<!-- LOAD DATA                                                           -->
<!-- Replace the path below with your own survey file.                   -->
<!-- The sample survey has 6 sheets: ROSTER, DWELLING, INCOME,           -->
<!-- FOOD_CONSUMPTION, HEALTH, EDUCATION                                 -->
<!-- ================================================================== -->

```{r load-data}
data_path <- "data/sample_survey.xlsx"

roster    <- read_excel(data_path, sheet = "ROSTER")
dwelling  <- read_excel(data_path, sheet = "DWELLING")
income    <- read_excel(data_path, sheet = "INCOME")
food      <- read_excel(data_path, sheet = "FOOD_CONSUMPTION")
health    <- read_excel(data_path, sheet = "HEALTH")
education <- read_excel(data_path, sheet = "EDUCATION")

roster <- roster %>% mutate(age_num = as.numeric(age))
```

\newpage

# Executive Summary

```{r summary-calcs}
n_dwelling_rows <- nrow(dwelling)
n_hh_unique     <- n_distinct(dwelling$hh_id)
n_persons       <- nrow(roster)

dup_hh    <- dwelling %>% group_by(hh_id) %>% filter(n() > 1)
n_dup_hh  <- n_distinct(dup_hh$hh_id)

heads_per_hh <- roster %>%
  filter(relationship == "head") %>%
  group_by(hh_id) %>%
  summarise(n_heads = n(), .groups = "drop")

hh_no_head  <- roster %>% distinct(hh_id) %>% anti_join(heads_per_hh, by = "hh_id")
n_no_head   <- nrow(hh_no_head)

has_critical <- (n_dup_hh > 0) | (n_no_head > 0)
```

| Concept | Count |
|:--------|:------|
| Dwelling records (rows) | `r format(n_dwelling_rows, big.mark=",")` |
| Unique households | `r format(n_hh_unique, big.mark=",")` |
| Persons (roster) | `r format(n_persons, big.mark=",")` |
| Avg. persons per household | `r round(n_persons / n_hh_unique, 2)` |

**Preliminary ruling:** `r ifelse(has_critical, "**REJECTED** — Critical incidences must be resolved before proceeding with analysis.", "**ACCEPTED WITH OBSERVATIONS** — The database can be used after minor corrections.")`

\newpage

# Duplicate Verification

## Household ID Duplicates (Dwelling Sheet)

Every household must appear exactly once in the dwelling sheet. Duplicates indicate a household was registered more than once.

```{r dup-hh}
if(n_dup_hh > 0) {
  add_incidence("critical", "Duplicates", "hh_id (dwelling)",
                paste(n_dup_hh, "duplicated hh_id"), nrow(dup_hh))
}
```

**Result:** `r ifelse(n_dup_hh == 0, "No duplicate household IDs found.", paste("**CRITICAL:**", n_dup_hh, "hh_id values appear more than once, affecting", nrow(dup_hh), "records."))`

```{r show-dups, results='asis'}
if(n_dup_hh > 0) {
  cat("\n**Duplicated hh_id:**\n\n")
  for(id in unique(dup_hh$hh_id)) {
    cat("- ", format(id, scientific = FALSE), "\n", sep="")
  }
}
```

## Person ID Duplicates (Roster)

```{r dup-person}
dup_person   <- roster %>% group_by(member_id) %>% filter(n() > 1)
n_dup_person <- n_distinct(dup_person$member_id)

if(n_dup_person > 0) {
  add_incidence("critical", "Duplicates", "member_id",
                paste(n_dup_person, "duplicated member_id"), nrow(dup_person))
}
```

**Result:** `r ifelse(n_dup_person == 0, "All persons have a unique identifier.", paste("**CRITICAL:**", n_dup_person, "duplicated member_id values."))`

\newpage

# Household Composition (Roster)

## Households Without Head

Every household must have exactly one head (relationship = "head").

```{r no-head}
if(n_no_head > 0) {
  add_incidence("critical", "Composition", "Head of household",
                paste(n_no_head, "households without head"), n_no_head)
}
```

**Result:** `r ifelse(n_no_head == 0, "All households have at least one head of household.", paste("**CRITICAL:**", n_no_head, "households have no head of household identified."))`

```{r show-no-head, results='asis'}
if(n_no_head > 0) {
  cat("\n**hh_id without head:**\n\n")
  for(id in hh_no_head$hh_id) {
    cat("- ", format(id, scientific = FALSE), "\n", sep="")
  }
}
```

## Multiple Heads

```{r multi-head}
multi_head   <- heads_per_hh %>% filter(n_heads > 1)
n_multi_head <- nrow(multi_head)

if(n_multi_head > 0) {
  add_incidence("medium", "Composition", "Multiple heads",
                paste(n_multi_head, "households with >1 head"), n_multi_head)
}
```

**Result:** `r ifelse(n_multi_head == 0, "All households have exactly one head.", paste("**MEDIUM:**", n_multi_head, "households have more than one head."))`

## Head of Household Under Age 15

```{r minor-head}
minor_heads   <- roster %>% filter(relationship == "head" & age_num < 15)
n_minor_heads <- nrow(minor_heads)

if(n_minor_heads > 0) {
  add_incidence("critical", "Composition", "Head under 15",
                paste(n_minor_heads, "heads under 15"), n_minor_heads)
}
```

**Result:** `r ifelse(n_minor_heads == 0, "All heads of household are 15 years or older.", paste("**CRITICAL:**", n_minor_heads, "heads of household are under 15 years old."))`

\newpage

# Dwelling Conditions — Missing Values

These variables have no skip conditions and must be complete for all households.

```{r dwelling-missing}
mandatory_dwelling <- c("wall_material", "roof_material", "floor_material", 
                        "n_rooms", "cooking_fuel", "main_water_source")

dwelling_result <- data.frame(Variable = character(), N_Missing = integer(), 
                              Result = character(), stringsAsFactors = FALSE)

for(v in mandatory_dwelling) {
  if(v %in% names(dwelling)) {
    n_miss <- sum(is.na(dwelling[[v]]) | dwelling[[v]] %in% c("", "NULL", " "))
    res <- ifelse(n_miss == 0, "OK", paste(n_miss, "missing"))
    dwelling_result <- rbind(dwelling_result, data.frame(Variable = v, N_Missing = n_miss, Result = res))
    if(n_miss > 0) {
      add_incidence("medium", "Dwelling", v, paste(n_miss, "missing values"), n_miss)
    }
  }
}

kable(dwelling_result, booktabs = TRUE, 
      caption = "Dwelling variables — missing values check") %>%
  kable_styling(latex_options = c("hold_position"))
```

\newpage

# Dwelling — Skip Logic Verification

```{r dwelling-skips}
n_error_rent <- NA
n_error_elec <- NA
n_error_toilet <- NA

# monthly_rent should only be answered if owns_dwelling = "no"
if("monthly_rent" %in% names(dwelling)) {
  error_rent <- dwelling %>%
    filter(owns_dwelling == "yes" & !is.na(monthly_rent))
  n_error_rent <- nrow(error_rent)
  if(n_error_rent > 0) {
    add_incidence("medium", "Dwelling", "monthly_rent (skip)",
                  paste(n_error_rent, "owners reported rent"), n_error_rent)
  }
}

# hours_electricity_day should only be answered if electricity_source != "none"
if("hours_electricity_day" %in% names(dwelling)) {
  error_elec <- dwelling %>%
    filter(electricity_source == "none" & !is.na(hours_electricity_day))
  n_error_elec <- nrow(error_elec)
  if(n_error_elec > 0) {
    add_incidence("medium", "Dwelling", "hours_electricity_day (skip)",
                  paste(n_error_elec, "households without electricity reported hours"), n_error_elec)
  }
}

# shared_toilet should only be answered if toilet_type != "none"
if("shared_toilet" %in% names(dwelling)) {
  error_toilet <- dwelling %>%
    filter(toilet_type == "none" & !is.na(shared_toilet))
  n_error_toilet <- nrow(error_toilet)
  if(n_error_toilet > 0) {
    add_incidence("medium", "Dwelling", "shared_toilet (skip)",
                  paste(n_error_toilet, "households without toilet reported sharing"), n_error_toilet)
  }
}
```

| Question | Skip condition | Improper responses | Result |
|:---------|:---------------|:-------------------|:-------|
| Monthly rent | Only if owns_dwelling = "no" | `r ifelse(is.na(n_error_rent), "N/A", n_error_rent)` | `r ifelse(is.na(n_error_rent), "Variable not found", ifelse(n_error_rent == 0, "OK", paste(n_error_rent, "errors")))` |
| Hours of electricity | Only if electricity_source ≠ "none" | `r ifelse(is.na(n_error_elec), "N/A", n_error_elec)` | `r ifelse(is.na(n_error_elec), "Variable not found", ifelse(n_error_elec == 0, "OK", paste(n_error_elec, "errors")))` |
| Shared toilet | Only if toilet_type ≠ "none" | `r ifelse(is.na(n_error_toilet), "N/A", n_error_toilet)` | `r ifelse(is.na(n_error_toilet), "Variable not found", ifelse(n_error_toilet == 0, "OK", paste(n_error_toilet, "errors")))` |

\newpage

# Income Module — Skip Logic Verification

```{r income-skips}
n_error_activity <- NA
n_error_remit <- NA

# main_activity should only be answered if worked_last_30_days = "yes"
if("main_activity" %in% names(income)) {
  error_activity <- income %>%
    filter(worked_last_30_days == "no" & !is.na(main_activity))
  n_error_activity <- nrow(error_activity)
  if(n_error_activity > 0) {
    add_incidence("medium", "Income", "main_activity (skip)",
                  paste(n_error_activity, "non-workers reported activity"), n_error_activity)
  }
}

# remittance_amount should only be answered if received_remittances = "yes"
if("remittance_amount" %in% names(income)) {
  error_remit <- income %>%
    filter(received_remittances == "no" & !is.na(remittance_amount))
  n_error_remit <- nrow(error_remit)
  if(n_error_remit > 0) {
    add_incidence("medium", "Income", "remittance_amount (skip)",
                  paste(n_error_remit, "non-recipients reported amount"), n_error_remit)
  }
}
```

| Question | Skip condition | Improper responses | Result |
|:---------|:---------------|:-------------------|:-------|
| Main activity | Only if worked_last_30_days = "yes" | `r ifelse(is.na(n_error_activity), "N/A", n_error_activity)` | `r ifelse(is.na(n_error_activity), "Variable not found", ifelse(n_error_activity == 0, "OK", paste(n_error_activity, "errors")))` |
| Remittance amount | Only if received_remittances = "yes" | `r ifelse(is.na(n_error_remit), "N/A", n_error_remit)` | `r ifelse(is.na(n_error_remit), "Variable not found", ifelse(n_error_remit == 0, "OK", paste(n_error_remit, "errors")))` |

\newpage

# Food Consumption — Skip Logic Verification

For each food group, days consumed and source should only be answered if the household reported consuming that food group.

```{r food-skips}
food_groups <- c("cereals", "roots_tubers", "legumes", "vegetables", "fruits", 
                 "meat", "fish", "dairy", "oils_fats", "sugar")

food_result <- data.frame(Food_Group = character(), N_Errors_Days = integer(), 
                          N_Errors_Source = integer(), stringsAsFactors = FALSE)

for(fg in food_groups) {
  consumed_var <- paste0("consumed_", fg)
  days_var     <- paste0("days_", fg)
  source_var   <- paste0("source_", fg)
  
  n_err_days <- 0
  n_err_source <- 0
  
  if(all(c(consumed_var, days_var) %in% names(food))) {
    n_err_days <- food %>%
      filter(.data[[consumed_var]] == "no" & !is.na(.data[[days_var]])) %>%
      nrow()
  }
  
  if(all(c(consumed_var, source_var) %in% names(food))) {
    n_err_source <- food %>%
      filter(.data[[consumed_var]] == "no" & !is.na(.data[[source_var]])) %>%
      nrow()
  }
  
  food_result <- rbind(food_result, data.frame(
    Food_Group = fg, N_Errors_Days = n_err_days, N_Errors_Source = n_err_source))
  
  if(n_err_days > 0) {
    add_incidence("medium", "Food consumption", paste0(days_var, " (skip)"),
                  paste(n_err_days, "non-consumers reported days"), n_err_days)
  }
  if(n_err_source > 0) {
    add_incidence("medium", "Food consumption", paste0(source_var, " (skip)"),
                  paste(n_err_source, "non-consumers reported source"), n_err_source)
  }
}

kable(food_result, booktabs = TRUE, col.names = c("Food group", "Days errors", "Source errors"),
      caption = "Food consumption module — skip logic errors by food group") %>%
  kable_styling(latex_options = c("hold_position"))
```

\newpage

# Health Module — Skip Logic Verification

```{r health-skips}
n_error_treatment <- NA
n_error_coverage <- NA

# sought_treatment should only be answered if illness_last_30_days = "yes"
if("sought_treatment" %in% names(health)) {
  error_treatment <- health %>%
    filter(illness_last_30_days == "no" & !is.na(sought_treatment))
  n_error_treatment <- nrow(error_treatment)
  if(n_error_treatment > 0) {
    add_incidence("medium", "Health", "sought_treatment (skip)",
                  paste(n_error_treatment, "non-ill persons reported treatment-seeking"), n_error_treatment)
  }
}

# coverage_type should only be answered if has_health_coverage = "yes"
if("coverage_type" %in% names(health)) {
  error_coverage <- health %>%
    filter(has_health_coverage == "no" & !is.na(coverage_type))
  n_error_coverage <- nrow(error_coverage)
  if(n_error_coverage > 0) {
    add_incidence("medium", "Health", "coverage_type (skip)",
                  paste(n_error_coverage, "uncovered persons reported coverage type"), n_error_coverage)
  }
}
```

| Question | Skip condition | Improper responses | Result |
|:---------|:---------------|:-------------------|:-------|
| Sought treatment | Only if illness_last_30_days = "yes" | `r ifelse(is.na(n_error_treatment), "N/A", n_error_treatment)` | `r ifelse(is.na(n_error_treatment), "Variable not found", ifelse(n_error_treatment == 0, "OK", paste(n_error_treatment, "errors")))` |
| Coverage type | Only if has_health_coverage = "yes" | `r ifelse(is.na(n_error_coverage), "N/A", n_error_coverage)` | `r ifelse(is.na(n_error_coverage), "Variable not found", ifelse(n_error_coverage == 0, "OK", paste(n_error_coverage, "errors")))` |

\newpage

# Education Module — Skip Logic Verification

```{r education-skips}
n_error_school <- NA

# school_type should only be answered if currently_enrolled = "yes"
if("school_type" %in% names(education)) {
  error_school <- education %>%
    filter(currently_enrolled == "no" & !is.na(school_type))
  n_error_school <- nrow(error_school)
  if(n_error_school > 0) {
    add_incidence("medium", "Education", "school_type (skip)",
                  paste(n_error_school, "non-enrolled persons reported school type"), n_error_school)
  }
}
```

**Result:** `r ifelse(is.na(n_error_school), "Variable not found.", ifelse(n_error_school == 0, "All records comply with the skip rule.", paste("**MEDIUM:**", n_error_school, "persons who are not currently enrolled reported a school type.")))`

\newpage

# Functional Screening — Missing Values (Roster)

Four functional domains must be answered for all persons in the roster.

```{r functional-check}
func_vars <- c("func_mobility", "func_cognition", "func_vision", "func_daily_tasks")
func_labels <- c("Mobility", "Cognition", "Vision", "Daily tasks")

func_result <- data.frame(Domain = character(), Variable = character(),
                          N_Missing = integer(), stringsAsFactors = FALSE)

for(i in seq_along(func_vars)) {
  var <- func_vars[i]
  if(var %in% names(roster)) {
    n_miss <- sum(is.na(roster[[var]]) | roster[[var]] %in% c("", "NULL", " "))
    func_result <- rbind(func_result,
                         data.frame(Domain = func_labels[i], Variable = var, N_Missing = n_miss))
    if(n_miss > 0) {
      add_incidence("medium", "Functional screening", func_labels[i],
                    paste(n_miss, "missing responses"), n_miss)
    }
  }
}

kable(func_result, booktabs = TRUE,
      caption = "Functional screening domains — missing values") %>%
  kable_styling(latex_options = c("hold_position"))
```

\newpage

# Incidence Summary

```{r incidence-summary}
n_critical <- nrow(incidences$critical)
n_medium   <- nrow(incidences$medium)
n_minor    <- nrow(incidences$minor)
```

| Severity | Count |
|:---------|:------|
| Critical | `r n_critical` |
| Medium | `r n_medium` |
| Minor | `r n_minor` |
| **Total** | **`r n_critical + n_medium + n_minor`** |

## Critical Incidences

```{r show-critical, results='asis'}
if(n_critical > 0) {
  print(kable(incidences$critical, booktabs = TRUE, 
              col.names = c("Section", "Variable", "Description", "N affected")) %>%
        kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9))
} else {
  cat("No critical incidences found.\n")
}
```

## Medium Incidences

```{r show-medium, results='asis'}
if(n_medium > 0) {
  print(kable(incidences$medium, booktabs = TRUE, 
              col.names = c("Section", "Variable", "Description", "N affected")) %>%
        kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9))
} else {
  cat("No medium incidences found.\n")
}
```

## Minor Incidences

```{r show-minor, results='asis'}
if(n_minor > 0) {
  print(kable(incidences$minor, booktabs = TRUE, 
              col.names = c("Section", "Variable", "Description", "N affected")) %>%
        kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9))
} else {
  cat("No minor incidences found.\n")
}
```

\newpage

# Ruling

```{r ruling}
ruling <- if(n_critical > 0) "REJECTED" 
          else if(n_medium > 0 | n_minor > 0) "ACCEPTED WITH OBSERVATIONS" 
          else "ACCEPTED"
```

## Ruling: `r ruling`

`r if(ruling == "REJECTED") "The database **cannot be used for analysis** in its current state due to the critical incidences identified above. The data provider must resolve these issues and resubmit the database. Re-run this document after corrections to verify."`

`r if(ruling == "ACCEPTED WITH OBSERVATIONS") "The database **can be used for analysis** after addressing the medium and minor incidences documented above. These issues do not prevent analysis but should be documented and corrected where possible."`

`r if(ruling == "ACCEPTED") "The database **is ready for analysis**. No incidences were found."`

---

*Document generated on `r Sys.Date()` — Re-run this .Rmd after any database correction to update all findings and the ruling automatically.*
